# Virus data analysis

##  Data load and restructure

Here I develop the code to import the data ... 

```{r}
library(maps)
library(maptools)

us_virus = read.csv(file = "./data/US_influenza.csv")

head(us_virus)

cols_of_interest = c('Subtype', 'Collection.Date', 'State.Province',
                     'Segment', 'Segment.Length')
us_virus = us_virus[ , cols_of_interest]

head(us_virus)

map('usa')
states = map('state', fill=T)
IDs = states$names
IDs_cr = sapply(IDs, function(x) strsplit(x, ":")[[1]][1])

state_sp = map2SpatialPolygons(states,IDs_cr, CRS("+proj=longlat"))

plot(state_sp)
points(coordinates(state_sp), col='red', pch='.')

state_coords = coordinates(state_sp)
rownames(state_coords)
```

## Setup Enviornmental Raster Layers. Graciously given to me from  Dr. McGlinn.

```{r}
library(raster)

prjString = "+proj=longlat +ellps=WGS84"

# data downloaded and described here
# http://www.worldclim.org/bioclim

# A function that corrects for the fact that WorldClim data is signed
fixData = function(x){
  temp = ifelse(x > 23767, x - 65536, x)
	new.x = ifelse(temp == -9999, NA, temp)
	new.x
}

# Stack WorldClim Bioclimatic Variables
useVars = 1:19
filenames = sapply(useVars, function(x) paste('bio', x, '.bil', sep=''))
tempStack = stack(file.path('./data/bioclim/', filenames))
bioStack = calc(tempStack, fixData)
projection(bioStack) = CRS(prjString)

# correct for the fact that the temperature related variables need
# to be devided by 10 to be in units of C
bioStack2 = bioStack
for(i in 1:11){  ##any temperature related variable
	bioStack2 = setValues(bioStack2, getValues(bioStack[[i]] / 10), i)
}
bioStack = bioStack2

names(bioStack) = c('mat','mdr','iso','tseas','tmax','tmin','tar','twetq',
                    'tdryq','twarmq','tcoldq','ap','pwet','pdry','pseas',
                    'pwetq','pdryq','pwarmq','pcoldq')

writeRaster(bioStack, file='./data/bioclim_5m.grd', format='raster',
            overwrite=TRUE)
```

## Query the enviornmental BioClim variables and extract state centroids

```{r}
bioStack = stack('./data/bioclim_5m.grd')
bioStack = crop(bioStack, extent(c(-135, -45, 22, 56)))
names(bioStack) = c('mat','mdr','iso','tseas','tmax','tmin','tar','twetq',
                    'tdryq','twarmq','tcoldq','ap','pwet','pdry','pseas',
                    'pwetq','pdryq','pwarmq','pcoldq')
# extract climate for entire state and average the values
state_clim = extract(bioStack, state_sp, fun=mean)
rownames(state_clim) = names(state_sp)
head(state_clim)

# first drop any rows in us_virus where State.Providice does not match
# a state name

us_virus = subset(us_virus, tolower(State.Province) %in% rownames(st_coords))

# now find the row indices of the state coord and climate matricies that 
# contain information on each record in the State.Provice record
row_indices = match(tolower(us_virus$State.Province), rownames(st_coords))
summary(row_indices)
length(row_indices)

coords = state_coords[row_indices, ]
clim = state_clim[row_indices, ]

us_virus = data.frame(us_virus, latitude = coords[ , 2],
                      longitude = coords[ , 1], clim)

head(us_virus)

write.csv(us_virus, file='./data/US_influenza_complete.csv', row.names=F)
```

##Building Models with the completed us_virus data set.

```{r}
#Package Download
#install.packages('dismo')
#install.packages('rJava')
library(dismo)

#load in the data.
us_virus_complete = read.csv(file = "./data/US_influenza_complete.csv")

#columns to keep for the linear model.
cols_to_keep = c('mat', 'tmax', 'tmin', 'Subtype','Collection.Date','State.Province')


##Building the models for specific subtypes, found in the data.
occ_h1n1 = us_virus_complete[us_virus_complete$Subtype == "H1N1", c("longitude", "latitude")]
occ_h3n2 = us_virus_complete[us_virus_complete$Subtype == "H3N2", c("longitude", "latitude")]
occ_h3 = us_virus_complete[us_virus_complete$Subtype == "H3", c("longitude", "latitude")]
occ_n1 = us_virus_complete[us_virus_complete$Subtype == "N1", c("longitude", "latitude")]

##Withholding 20% sample. 
#H1N1
fold_h1n1 <- kfold(occ_h1n1, k=5)
occtest_h1n1 <- occ_h1n1[fold_h1n1 == 1, ]
occtrain_h1n1 <- occ_h1n1[fold_h1n1 != 1, ]
#H3N2
fold_h3n2 <- kfold(occ_h3n2, k=5)
occtest_h3n2 <- occ_h1n1[fold_h3n2 == 1, ]
occtrain_h3n2 <- occ_h1n1[fold_h3n2 != 1, ]
#N1
fold_n1 <- kfold(occ_n1, k=5)
occtest_n1 <- occ_h1n1[fold_n1 == 1, ]
occtrain_n1 <- occ_h1n1[fold_n1 != 1, ]
#H3
fold_h3 <- kfold(occ_h3, k=5)
occtest_h3 <- occ_h1n1[fold_h3 == 1, ]
occtrain_h3 <- occ_h1n1[fold_h3 != 1, ]

## Maxent Models
me_h1n1 = maxent(bioStack, occtrain_h1n1)
me_h3n2 = maxent(bioStack, occtrain_h3n2)
me_h3 = maxent(bioStack, occtrain_h3)
me_n1 = maxent(bioStack, occtrain_n1)


